\modLP{Identification of} Biological Regulatory Networks from Process Hitting models

### 2014-11-03 parts-phage-lambda.tex
examples++

  % Sortes
  \TSetSortLabel{cI}{$\cI$}
  \TSort{(0,3.5)}{cI}{3}{l}
  \TSetSortLabel{cro}{$\cro$}
  \TSort{(0,-1.5)}{cro}{4}{l}
  \TSetSortLabel{N}{$\N$}
  \TSort{(7.5,1.5)}{N}{2}{r}

@@

  % Actions de MàJ
  \THit{cI_1}{}{cIcro_0}{.west}{cIcro_2}
  \THit{cI_1}{}{cIcro_1}{.160}{cIcro_3}
  \THit{cI_2}{}{cIcro_0}{.west}{cIcro_2}
  \THit{cI_2}{}{cIcro_1}{.160}{cIcro_3}
  \path[bounce,bend left]
    \TBounce{cIcro_0}{}{cIcro_2}{.240}
    \TBounce{cIcro_1}{}{cIcro_3}{.south west} ;
  \THit{cI_0}{}{cIcro_2}{.160}{cIcro_0}
  \THit{cI_0}{}{cIcro_3}{.150}{cIcro_1}
  \path[bounce,bend right]
    \TBounce{cIcro_2}{}{cIcro_0}{.north west}
    \TBounce{cIcro_3}{}{cIcro_1}{.120} ;
  \THit{cro_2}{}{cIcro_0}{.210}{cIcro_1}
  \THit{cro_2}{}{cIcro_2}{.200}{cIcro_3}
  \THit{cro_3}{}{cIcro_0}{.210}{cIcro_1}
  \THit{cro_3}{}{cIcro_2}{.200}{cIcro_3}
  \path[bounce,bend left]
    \TBounce{cIcro_0}{}{cIcro_1}{.240}
    \TBounce{cIcro_2}{}{cIcro_3}{.south} ;
  \THit{cro_0}{}{cIcro_3}{.190}{cIcro_2}
  \THit{cro_0}{}{cIcro_1}{.200}{cIcro_0}
  \THit{cro_1}{}{cIcro_3}{.190}{cIcro_2}
  \THit{cro_1}{}{cIcro_1}{.200}{cIcro_0}
  \path[bounce,bend right]
    \TBounce{cIcro_1}{}{cIcro_0}{.north}
    \TBounce{cIcro_3}{}{cIcro_2}{.120} ;

  % Actions sortantes
  \THit{cIcro_3}{}{N_1}{.west}{N_0}
  \path[bounce,bend right] \TBounce{N_1}{}{N_0}{.north} ;
  \THit{cIcro_2}{}{N_0}{.north west}{N_1}
  \path[bounce,bend left] \TBounce{N_0}{}{N_1}{.260} ;
  \THit{cIcro_1}{}{N_0}{.west}{N_1}
  \path[bounce,bend left] \TBounce{N_0}{}{N_1}{.230} ;
  \THit{cIcro_0}{}{N_0}{.south west}{N_1}
  \path[bounce,bend left=40] \TBounce{N_0}{}{N_1}{.200} ;
\end{tikzpicture}
}
\caption{\label{fig:phage-lambda-cIcro-N}%
  A portion of the PH model used in this section modeling the phage lambda immunity response.
  This part of the whole model only depicts the interactions towards $\N$ from components
  $\cI$ and $\cro$.
  The processes of the cooperative sort $\cIcro$ are labeled by the configuration they represent,
  given the values of $\cI$ and $\cro$ regarding their respective thresholds.
}
\end{figure}

We first note that in the case where $\cro_3$ is present, we have from \pref{eq:ctx-sigma}:
\begin{align*}
  \ctx_\N^{\{ \cI, \cro \}}(\state{\cI_0, \cro_3, \N_0})
    &= \{ \cI_0, \cro_3, \N_0, \cIcro_{FT} \} \\
  \ctx_\N^{\{ \cI, \cro \}}(\state{\cI_1, \cro_3, \N_0})
    &= \{ \cI_1, \cro_3, \N_0, \cIcro_{TT} \} \\
  \ctx_\N^{\{ \cI, \cro \}}(\state{\cI_2, \cro_3, \N_0})
    &= \{ \cI_2, \cro_3, \N_0, \cIcro_{TT} \}
\end{align*}
We can then deduce the following directions of evolution from \pref{eq:irB}:
\begin{align*}
  \irB_\N^{\{ \cI, \cro \}}(\state{\cI_0, \cro_3, \N_0}) &= \{ \N_1 \} \\
%     \text{because } \state{\cI_0, \cro_3, \N_0} \in \Phi(\N_0) \\
%     \text{(from \pref{eq:B-no-bounce})} \\
  \irB_\N^{\{ \cI, \cro \}}(\state{\cI_1, \cro_3, \N_0}) &= \{ \N_0 \} \\
%     \text{because } \state{\cI_0, \cro_3, \N_0} \in \Phi(\N_0) \\
%     \text{(from \pref{eq:B-bounce})} \\
  \irB_\N^{\{ \cI, \cro \}}(\state{\cI_2, \cro_3, \N_0}) &= \{ \N_0 \}
%     \text{because } \state{\cI_0, \cro_3, \N_0} \in \Phi(\N_0) \\
%     \text{(from \pref{eq:B-bounce})}
\end{align*}
          %      eq:inference-edges-ba
Therefore, \pref{eq:edges-inference} gives: $\cI \xrightarrow{1} \N \in \hat{E}_-$.
The configurations using the other processes of $\cro$ ($\cro_0$, $\cro_1$ and $\cro_2$)
do not add more information on the influences towards $\N$.
Thus, given \pref{pps:inference-IG}, it comes: $\GRNedgef{\cI}{-}{1}{\N} \in E$.

@@

Let us first focus on the group of regulators $\{ \cro \}$.
Of course:
\[\forall \cro_i \in \PHl_{\cro}, \ctx_{\cro}^{\{ \cro \}}(\state{\cro_i}) = \{ \cro_i \}\]
and from \pref{eq:irB} and \eqref{eq:fp}:
\begin{align*}
  \irB_{\cro}^{\{ \cro \}}(\state{\cro_3}) &= \{ \cro_2 \} && \\
%     \text{(from \pref{eq:B-bounce})} \\
  \irB_{\cro}^{\{ \cro \}}(\state{\cro_2}) &= \{ \cro_1 \} && \text{and} &
    \Phi&(\cro_2) = \emptyset \\
%     \text{(from \pref{eq:B-no-bounce})} \\
  \irB_{\cro}^{\{ \cro \}}(\state{\cro_1}) &= \{ \cro_1 \} && \text{and} &
    \Phi&(\cro_1) = \emptyset \\
%     \text{(from \pref{eq:B-no-bounce})} \\
  \irB_{\cro}^{\{ \cro \}}(\state{\cro_0}) &= \{ \cro_0 \} && \text{and} &
    \state{\cI_2, \cro_0} \in \Phi&(\cro_0)
%     \text{(from \pref{eq:B-no-bounce})}
\end{align*}
We thus cannot infer any local influence from \pref{eq:edges-inference-auto}
by observing the group of regulators $\{ \cro \}$,
because $\Phi(\cro_1) = \Phi(\cro_2) = \emptyset$.
This is due to the fact that all processes of $\cI$ can hit $\cro_1$ and $\cro_2$;
these processes can therefore not be considered stable in any configuration.

Let us now focus instead on the group of regulators $\{ \cI \}$.
One again, we trivially have:
\[\forall \cI_i \in \PHl_{\cI}, \forall \cro_j \in \PHl_{\cro},
  \ctx_{\cro}^{\{ \cI \}}(\state{\cI_i, \cro_j}) = \{ \cI_i, \cro_j \}\]
Furthermore, from \pref{eq:irB}, we especially find:
\begin{align*}
%   \irB_{\cro}^{\{ \cI, \cro \}}(\state{\cI_0, \cro_0}) &= \{ \cro_1 \} &&
%   \text{(from \pref{eq:B-bounce})} \\
%   \irB_{\cro}^{\{ \cI, \cro \}}(\state{\cI_0, \cro_1}) &= \{ \cro_2 \} &&
%   \text{(from \pref{eq:B-bounce})} \\
  \irB_{\cro}^{\{ \cI \}}(\state{\cI_0, \cro_2}) &= \{ \cro_3 \} \\
%     \text{(from \pref{eq:B-bounce})} \\
  \irB_{\cro}^{\{ \cI \}}(\state{\cI_0, \cro_3}) &= \{ \cro_2 \}
%     \text{(from \pref{eq:B-bounce})}
\end{align*}
Therefore, \pref{eq:edges-inference} gives: $\cro \xrightarrow{3} \cro \in \hat{E}_-$.
% The configurations using the other processes of $\cro$ ($\cro_0$, $\cro_1$ and $\cro_2$)
% do not add more information on the influences towards $\N$.
Finally, as no more information is added by observing the other levels of $\cro$ and $\cI$,
and given \pref{pps:inference-IG}, it comes: $\GRNedgef{\cro}{-}{3}{\cro} \in E$.

@@

\begin{table}[t]
\begin{align*}
  K_{\cI, \emptyset} &= 2 &
  K_{\N, \emptyset} &= 1\\
  K_{\cI, \{\cro\}} &= 0&
  K_{\N, \{\cro\}} &= 0\\
  K_{\cI, \{\cII\}} &= 2&
  K_{\N, \{\cI\}} &= 0\\
  K_{\cI, \{\cII;cro\}} &= 2&
  K_{\N, \{\cI;cro\}} &= 0\\
  \\
  K_{\cII, \emptyset} &= 0&
  K_{\cro, \emptyset} &= 3\\
  K_{\cII, \{\cro\}} &= 0&
  K_{\cro, \{\cro\}} &= 2\\
  K_{\cII, \{\cI\}} &= 0&
  K_{\cro, \{\cI\}} &= 0\\
  K_{\cII, \{\cI;cro\}} &= 0&
  K_{\cro, \{\cI;cro\}} &= 2\\
  K_{\cII, \{\N\}} &= 1\\
  K_{\cII, \{\N;cro\}} &= 0\\
  K_{\cII, \{\N;cI\}} &= 0\\
  K_{\cII, \{\N;cI;cro\}} &= 0
\end{align*}
\caption{\label{tb:phage-lambda-k}%
  Result of the parameters inference performed on the Process Hitting model
  of the phage lambda immunity response
  taken from \cite{thieffry_dynamical_1995}.
}
\end{table}

We note two main differences between the parametrization obtained by our inference
and the one proposed in \cite{thieffry_dynamical_1995}:
\begin{itemize}
  \item The absence of the self-regulation of $\cI$ automatically reduces the size
    of the parametrization, but this does not modify the dynamics, as mentioned previously;
  \item Our inference returns $K_{\cro, \{\cI;cro\}} = 2$ instead of originally $0$;
    once again, this does not change the dynamics, and $1$ would also have been
    an acceptable value.
\end{itemize}

>>>>

  % Sortes
  \TSetSortLabel{cI}{$\cI$}
  \TSort{(1,3.5)}{cI}{3}{l}
  \TSetSortLabel{cro}{$\cro$}
  \TSort{(1,-1.5)}{cro}{4}{l}
  \TSetSortLabel{N}{$\N$}
  \TSort{(7.5,1.5)}{N}{2}{r}

@@

  % Actions de MàJ
%   \THit{cI_1}{}{cIcro_0}{.west}{cIcro_2}
%   \THit{cI_1}{}{cIcro_1}{.160}{cIcro_3}
%   \THit{cI_2}{}{cIcro_0}{.west}{cIcro_2}
%   \THit{cI_2}{}{cIcro_1}{.160}{cIcro_3}
%   \path[bounce,bend left]
%     \TBounce{cIcro_0}{}{cIcro_2}{.240}
%     \TBounce{cIcro_1}{}{cIcro_3}{.south west} ;
%   \THit{cI_0}{}{cIcro_2}{.160}{cIcro_0}
%   \THit{cI_0}{}{cIcro_3}{.150}{cIcro_1}
%   \path[bounce,bend right]
%     \TBounce{cIcro_2}{}{cIcro_0}{.north west}
%     \TBounce{cIcro_3}{}{cIcro_1}{.120} ;
%   \THit{cro_2}{}{cIcro_0}{.210}{cIcro_1}
%   \THit{cro_2}{}{cIcro_2}{.200}{cIcro_3}
%   \THit{cro_3}{}{cIcro_0}{.210}{cIcro_1}
%   \THit{cro_3}{}{cIcro_2}{.200}{cIcro_3}
%   \path[bounce,bend left]
%     \TBounce{cIcro_0}{}{cIcro_1}{.240}
%     \TBounce{cIcro_2}{}{cIcro_3}{.south} ;
%   \THit{cro_0}{}{cIcro_3}{.190}{cIcro_2}
%   \THit{cro_0}{}{cIcro_1}{.200}{cIcro_0}
%   \THit{cro_1}{}{cIcro_3}{.190}{cIcro_2}
%   \THit{cro_1}{}{cIcro_1}{.200}{cIcro_0}
%   \path[bounce,bend right]
%     \TBounce{cIcro_1}{}{cIcro_0}{.north}
%     \TBounce{cIcro_3}{}{cIcro_2}{.120} ;

  % MàJ abstraites de la sorte coopératiove
  \path (1.8, 4.5) edge[coopupdate] (3.2, 3);
  \path (1.8, 0) edge[coopupdate] (3.2, 1);

  % Actions sortantes
  \THit{cIcro_3}{bend right=5}{N_1}{.150}{N_0}
  \path[bounce,bend right=80,in=220] \TBounce{N_1}{}{N_0}{.130} ;
  \THit{cIcro_2}{bend right=4}{N_0}{.150}{N_1}
  \path[bounce,bend left] \TBounce{N_0}{}{N_1}{.260} ;
  \THit{cIcro_1}{bend right=20}{N_0}{.west}{N_1}
  \path[bounce,bend left] \TBounce{N_0}{}{N_1}{.230} ;
  \THit{cIcro_0}{bend left=2}{N_0}{.south west}{N_1}
  \path[bounce,bend left=40] \TBounce{N_0}{}{N_1}{.200} ;
\end{tikzpicture}
}
\caption{\label{fig:phage-lambda-cIcro-N}%
  A portion of the PH model used in this section modeling the phage lambda immunity response.
  This part of the whole model only depicts the interactions towards $\N$ from components
  $\cI$ and $\cro$.
  All the actions updating the cooperative $\cIcro$ are simply depicted by two
  zigzag arrows.
  %, which abstract the fact that this cooperative sort is deterministic.
  Furthermore, as this cooperative sort $\cIcro$ is deterministic,
  its processes are labeled by the configuration they represent,
  given by two Boolean variables indicating whether
  the values of $\cI$ and $\cro$ lie above their respective thresholds ($T$) or not ($F$).
  For example, the configuration $\state{\cI_1, \cro_0}$ is represented by the process $TF$.
}
\end{figure}

We first note that in the case where $\cro_3$ is present, we have from \pref{eq:ctx-sigma}:
\begin{align*}
  \ctx_\N^{\{ \cI, \cro \}}(\state{\cI_0, \cro_3, \N_0})
    &= \state{\cI_0, \cro_3, \N_0, \cIcro_{FT}} \\
  \ctx_\N^{\{ \cI, \cro \}}(\state{\cI_1, \cro_3, \N_0})
    &= \state{\cI_1, \cro_3, \N_0, \cIcro_{TT}} \\
  \ctx_\N^{\{ \cI, \cro \}}(\state{\cI_2, \cro_3, \N_0})
    &= \state{\cI_2, \cro_3, \N_0, \cIcro_{TT}}
\end{align*}
We can then deduce the following directions of evolution from \pref{eq:irB}:
\begin{align*}
  \irB_\N^{\{ \cI, \cro \}}(\state{\cI_0, \cro_3, \N_0}) &= \{ \N_1 \} \\
%     \text{because } \state{\cI_0, \cro_3, \N_0} \in \Phi(\N_0) \\
%     \text{(from \pref{eq:B-no-bounce})} \\
  \irB_\N^{\{ \cI, \cro \}}(\state{\cI_1, \cro_3, \N_0}) &= \{ \N_0 \} \\
%     \text{because } \state{\cI_0, \cro_3, \N_0} \in \Phi(\N_0) \\
%     \text{(from \pref{eq:B-bounce})} \\
  \irB_\N^{\{ \cI, \cro \}}(\state{\cI_2, \cro_3, \N_0}) &= \{ \N_0 \}
%     \text{because } \state{\cI_0, \cro_3, \N_0} \in \Phi(\N_0) \\
%     \text{(from \pref{eq:B-bounce})}
\end{align*}
          %      eq:inference-edges-ba
Therefore, \pref{eq:edges-inference} gives: $\cI \xrightarrow{1} \N \in \hat{E}_-$.
The configurations using the other processes of $\cro$ ($\cro_0$, $\cro_1$ and $\cro_2$)
do not add more information on the influences towards $\N$.
Thus, given \pref{pps:inference-IG}, it comes: $\GRNedgef{\cI}{-}{1}{\N} \in E$.

@@

Let us first focus on the group of regulators $\{ \cro \}$.
Of course:
\[\forall \cro_i \in \PHl_{\cro}, \ctx_{\cro}^{\{ \cro \}}(\state{\cro_i}) = \state{\cro_i}\]
and from \pref{eq:irB} and \eqref{eq:fp}:
\begin{align*}
  \irB_{\cro}^{\{ \cro \}}(\state{\cro_3}) &= \{ \cro_2 \} && \\
%     \text{(from \pref{eq:B-bounce})} \\
  \irB_{\cro}^{\{ \cro \}}(\state{\cro_2}) &= \{ \cro_1 \} && \text{and} &
    \Phi&(\cro_2) = \emptyset \\
%     \text{(from \pref{eq:B-no-bounce})} \\
  \irB_{\cro}^{\{ \cro \}}(\state{\cro_1}) &= \{ \cro_1 \} && \text{and} &
    \Phi&(\cro_1) = \emptyset \\
%     \text{(from \pref{eq:B-no-bounce})} \\
  \irB_{\cro}^{\{ \cro \}}(\state{\cro_0}) &= \{ \cro_0 \} && \text{and} &
    \state{\cI_2, \cro_0} \in \Phi&(\cro_0)
%     \text{(from \pref{eq:B-no-bounce})}
\end{align*}
We thus cannot infer any local influence from \pref{eq:edges-inference-auto}
by observing the group of regulators $\{ \cro \}$,
because $\Phi(\cro_1) = \Phi(\cro_2) = \emptyset$.
This is due to the fact that all processes of $\cI$ can hit $\cro_1$ and $\cro_2$;
these processes can therefore not be considered stable in any configuration.

Let us now focus instead on the group of regulators $\{ \cI \}$.
One again, we trivially have:
\[\forall \cI_i \in \PHl_{\cI}, \forall \cro_j \in \PHl_{\cro},
  \ctx_{\cro}^{\{ \cI \}}(\state{\cI_i, \cro_j}) = \state{\cI_i, \cro_j}\]
Furthermore, from \pref{eq:irB}, we especially find:
\begin{align*}
%   \irB_{\cro}^{\{ \cI, \cro \}}(\state{\cI_0, \cro_0}) &= \{ \cro_1 \} &&
%   \text{(from \pref{eq:B-bounce})} \\
%   \irB_{\cro}^{\{ \cI, \cro \}}(\state{\cI_0, \cro_1}) &= \{ \cro_2 \} &&
%   \text{(from \pref{eq:B-bounce})} \\
  \irB_{\cro}^{\{ \cI \}}(\state{\cI_0, \cro_2}) &= \{ \cro_3 \} \\
%     \text{(from \pref{eq:B-bounce})} \\
  \irB_{\cro}^{\{ \cI \}}(\state{\cI_0, \cro_3}) &= \{ \cro_2 \}
%     \text{(from \pref{eq:B-bounce})}
\end{align*}
Therefore, \pref{eq:edges-inference} gives: $\cro \xrightarrow{3} \cro \in \hat{E}_-$.
% The configurations using the other processes of $\cro$ ($\cro_0$, $\cro_1$ and $\cro_2$)
% do not add more information on the influences towards $\N$.
Finally, as no more information is added by observing the other levels of $\cro$ and $\cI$,
and given \pref{pps:inference-IG}, it comes: $\GRNedgef{\cro}{-}{3}{\cro} \in E$.

@@

\begin{table}[t]
\begin{align*}
  K_{\cI, \emptyset} &= 2 &
  K_{\N, \emptyset} &= 1\\
  K_{\cI, \{\cro\}} &= 0&
  K_{\N, \{\cro\}} &= 0\\
  K_{\cI, \{\cII\}} &= 2&
  K_{\N, \{\cI\}} &= 0\\
  K_{\cI, \{\cII;\cro\}} &= 2&
  K_{\N, \{\cI;\cro\}} &= 0\\
  \\
  K_{\cII, \emptyset} &= 0&
  K_{\cro, \emptyset} &= 3\\
  K_{\cII, \{\cro\}} &= 0&
  K_{\cro, \{\cro\}} &= 2\\
  K_{\cII, \{\cI\}} &= 0&
  K_{\cro, \{\cI\}} &= 0\\
  K_{\cII, \{\cI;\cro\}} &= 0&
  K_{\cro, \{\cI;\cro\}} &= 2\\
  K_{\cII, \{\N\}} &= 1\\
  K_{\cII, \{\N;\cro\}} &= 0\\
  K_{\cII, \{\N;cI\}} &= 0\\
  K_{\cII, \{\N;cI;\cro\}} &= 0
\end{align*}
\caption{\label{tb:phage-lambda-k}%
  Result of the parameters inference performed on the Process Hitting model
  of the phage lambda immunity response
  taken from \cite{thieffry_dynamical_1995}.
}
\end{table}

We note two main differences between the parametrization obtained by our inference
and the one proposed in \cite{thieffry_dynamical_1995}:
\begin{itemize}
  \item The absence of the self-regulation of $\cI$ automatically reduces the size
    of the parametrization, but this does not modify the dynamics, as mentioned previously;
  \item Our inference returns $K_{\cro, \{\cI;\cro\}} = 2$, \modLP{which is a value
    equivalent to the original $0$ value according to \pref{rem:K-equiv}:
    the resulting dynamics is the same, and $1$ would also have been an
    acceptable value}.
\end{itemize}
